---
title: "EXPLORATORY DATA ANALYSIS"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Reading Bigger Data Files using read.csv() Command.
```{r}
d2 = read.csv(file.choose())
d2
```

Getting information of an Object using summary() command.
```{r}

summary(d2)

```

Use of names() and dimnames() command to view the column and row elements of an Object.
```{r}
rownames(d2)
colnames(d2)
plot(d2)
```
Summarizing samples: 


```{r}
list(d2)
length(d2)
head(d2)
tail(d2)
str(d2)
d2$Sp_defence
d2$d2

cbind(d2)
dimnames(d2)
names(d2)
mean(d2$Attack)
max(d2$HP)
quantile(d2$Sp_attack)
fivenum(d2$Defence)
```

Cumulative statistics:
```{r}
cumsum(d2$HP)
head(cumsum(d2$Total))
head(cumsum(d2$Attack))
head(cumsum(d2$Defence))
head(cumsum(d2$Sp_attack))
head(cumsum(d2$Sp_defence))
head(cumsum(d2$Speed))


cummax(d2$HP)
head(cummax(d2$Total))
head(cummax(d2$Attack))
head(cummax(d2$Defence))
head(cummax(d2$Sp_attack))
head(cummax(d2$Sp_defence))
head(cummax(d2$Speed))

cummin(d2$HP)
cummin(d2$Attack)
cummin(d2$Defence)
cummin(d2$Sp_attack)
cumprod(d2$HP)
seq_along(d2)
seq(from = 1, to = 5, by =2)
cumsum(d2$HP) / seq_along(d2) 
md = seq_along(d2)
md
mean(d2[,2])
mean(d2[2,])

table(d2$HP)
d.mat = as.matrix(d2)
d.mat

plot(d2$HP)
plot(d2$Total)
plot(d2$Attack)
plot(d2$Defence)
plot(d2$Sp_attack)
plot(d2$Sp_defence)
plot(d2$Speed)

plot(md)



stem(d2$HP)
stem(d2$Total)
stem(d2$Attack)
stem(d2$Defence)
stem(d2$Sp_attack)
stem(d2$Sp_defence)
stem(d2$Speed)
stem(d2$HP, scale =2)

hist(d2$HP)
hist(d2$Total)
hist(d2$Attack)
hist(d2$Defence)
hist(d2$Sp_attack)
hist(d2$Sp_defence)
hist(d2$Speed)
hist(d2$HP, breaks = 'Sturges')
hist(d2$HP, breaks = 'Scott')
hist(d2$HP, breaks = 'FD')
hist(d2$HP, breaks = 'st')
```


```{r}
hist(d2$HP, breaks = 'sc')
hist(d2$HP, breaks = 7)
hist(d2$HP, col = 'gray75', main = NULL, xlab = 'Size class of d', ylim = c(0,0.1), freq = FALSE)
dens = density(d2$HP)
dens
```

The class() command to know the type of object.

```{r}
class(d2)
class(dimnames(d2))

class(d2$HP)
```

The str() command to examine the structure of an object.

```{r}


str(d2)
str(d2$HP)
structure(head(d2))
ls.str(pattern = 'd2')

attach(head(d2))
stack(head(d2))

dimnames(d2)[[1]]
dimnames(d2)[[2]]
head(d2, n=10)

```

Simple math:

```{r}
search()
sin(d2$HP)
cos(head(d2$HP))
factorial(head(d2$HP))
abs(head(d2$HP))
sqrt(head(d2$HP))
tan(head(d2$HP))
asin(head(d2$HP))
```

Viewing Previously Loaded Named-Objects.

```{r}

ls()
ls(pattern = 'v')
ls(pattern = 'da')


```


```{r}
var(head(d2$HP))
median(d2$HP)
mode(d2$HP)
```

Selecting and Displaying Parts of a Object.

```{r}
head(d2[1])
head(d2[1:4])
head(d2[-1])
head(d2[d2>7])
head(d2[d2<5 | d2>10],n=10)
head(d2[d2<5 | d2>10])

d2[100,8]
d2[150,1:6]
```

Sorting and Rearranging a Object.

```{r}

sort(head(d2$HP))
sort(head(d2$HP), decreasing = TRUE)
sort(head(d2$HP),na.last = NA)
sort(head(d2$HP),na.last = TRUE)
sort(head(d2$HP),na.last = FALSE)

order(d2$HP)
rank(d2)
rank(d2$HP)

```


```{r}
head(d2[5])
with(d2, sum(Defence))
sd(d2$HP)
```

Summarizing rows and columns:
```{r}
mad(d2$Defence)
apply(head(d2),1, mean, na.rm=TRUE)
apply(head(d2),1, median, na.rm=TRUE)
is.table(d2)
addmargins(head(d2),1, mean)
addmargins(head(d2),1, median)
View(head(d2))
nrow((d2))
ncol(d2)
sum(is.na(d2))

structure(head(d2))
table(head((d2$Speed)))
plot(density(d2$HP))
```

Drawing distribution and Graphics:
```{r}
hist(d2$HP, freq = F, col= 'blue')
lines(density(d2$HP), lty= 2)
lines(density(d2$HP, k='rectangular'))
rnorm(20, mean=5, sd= 1)
pnorm(20, mean=5, sd= 1)
qnorm(0.5, 5, 1)
dnorm(c(7,8,9),mean=5, sd=1)
qnorm(c(3.5,0.75),mean=5, sd=1)
d2.norm=rnorm(1000,mean(d2$HP), sd(d2$HP))
head(d2.norm)
hist(d2$HP, freq = F, col= 'pink')
lines(density(d2.norm))
hist(d2.norm, freq = F, col= 'pink')
lines(density(d2$HP))
hist(d2.norm,freq = F,border = 'red',col='blue',main = 'comparing two distributions',xlab = 'data size classes')
lines(density(d2$HP), lwd= 2)
qqnorm(d2$HP)
qqnorm(d2$HP, main = 'QQ plot of example data', xlab = 'Theoretical',ylab = 'Quantiles for d2')
qqline(d2$HP, lwd = 2, lty = 2)
qqp = qqplot(d2$HP, rnorm(50,5,2))
qqp
abline(lm(qqp$y ~ qqp$x))
```

Testing distribution:
```{r}
sample(head(d2),size = 4)
set.seed(4)
sample(head(df), size = 3)
shapiro.test(d2$HP)
ks.test(d2$HP, 'pnorm',mean=5,sd=2)
ks.test(d2$HP, pnorm(20,5,2))
```

T-test:
```{r}
t.test(d2$HP, d2$Attack)
t.test(d2$HP, d2$Attack, var.equal = TRUE)
t.test(d2$HP, mu = 10)
t.test(d2$HP, mu = 1000, alternative = 'greater')
```

U-test:
```{r}
wilcox.test(d2$HP, d2$Attack)
wilcox.test(d2$HP, exact = FALSE)
wilcox.test(d2$HP, mu = 8, exact = FALSE, conf.int = TRUE, alt = 'less')
```

Correlation and Covariance:
```{r}
cor(d2$Defence, d2$Attack)
cor(d2$Attack, d2$Defence, method = 'spearman')
cor(d2$Attack, d2$Defence, method = 'kendall')
cov(d2$Attack, d2$Defence)
```


Correlation hypothesis tests:
```{r}
cor.test(d2$Attack, d2$Defence)
```

Association tests:
```{r}
chisq.test(d2$Attack)
chisq.test(d2$HP, correct = FALSE)
d2.cs = chisq.test(d2$Attack, p = d2$Defence, rescale.p = TRUE)
d2.cs
head(d2.cs$exp)
names(d2.cs$expected) = row.names(d2)
head(d2.cs$exp)
(honey.cs = chisq.test(d2$Attack))
names(honey.cs)
(honey.cs = chisq.test(d2$Attack, simulate.p.value = TRUE, B = 3000))
d2[1:2, 4:5]
chisq.test(d2[1:2, 4:5], correct = FALSE)
chisq.test(d2[1:2, 4:5], correct = TRUE)
with(d2, chisq.test(Attack, p = Defence, rescale = T))
with(d2, chisq.test(Attack, p =Defence , rescale = T, sim = T))
chisq.test(d2$Attack)
```

Box-whisker plots:

```{r}
boxplot(d2$Attack)
boxplot(d2$Attack, d2$Defence)
boxplot(d2$Attack, d2$Defence, names = c('Attack', 'Defence'))
title(xlab = 'Variable', ylab = 'Value')
boxplot(d2$Attack, d2$Defence, names = c('Attack', 'Defence'), range = 0,
xlab = 'Variable', ylab = 'Value', col = 'yellow')
boxplot(d2$Attack, d2$Defence, d2$HP, names = c('Attack', 'Defence', 'HP'),
range = 0, xlab = 'variable', ylab = 'value', col = 'pink')
```

Scatter plots:

```{r}
plot(d2$Attack, d2$Defence, pch = "+")

plot(d2$Attack, d2$Defence, xlab = 'Attack', ylab = 'Defence', pch = 10, cex = 2, col = 'blue', xlim = c(0, 500), ylim = c(0, 500))

plot(Defence ~ Attack, data = d2)
abline(lm(Defence ~ Attack, data = d2))

plot(Defence ~ Attack, data = d2)
abline(lm(d2$Attack ~ d2$Defence, data = d2), lty = 'dotted', lwd = 2, col = 'green')

plot(d2$Attack ~ d2$Defence, data = d2, xlab = 'Attack', ylab = 'Defence',
pch = 10, cex = 2, col = 'gray50', xlim = c(0, 500), ylim = c(0, 500))
abline(lm(d2$Attack ~ d2$Defence, data = d2), lty = 'dotted', lwd = 2, col = 'blue')

plot(d2$Attack, type = 'l')

with(d2, plot(Attack, HP, type = 'c'))
with(d2[order(d2$Attack),], plot(sort(Attack), HP, type = 'c'))

plot(d2$Attack, type = 'b')
```

Multiple correlation plots:

```{r}
pairs(~ Defence + Attack + HP, data = d2)

pairs(~ d2$Attack + d2$HP + d2$Defence, data = d2, col ='brown', cex = 5, pch = 'X')
```

Pie charts:
```{r}
pie(d2$HP, labels = d2$Total)
 
 pc = c('gray40', 'gray50', 'gray60', 'gray70', 'gray80', 'gray90')
 pie(head(d2$HP, labels = d2$Attack, col = pc, clockwise = TRUE, init.angle = 180))
 
 pc = c('gray65', 'gray70', 'gray75', 'gray80', 'gray85', 'gray90')
 pie(head(d2$Attack, labels = row.names(d2), col = pc, cex = 1.2))
 
 d2[1,]
```


Cleveland dot charts:

```{r}
dotchart(d2$Attack)
 dotchart(t(d2$HP))
 
 #Bar charts:
```

Bar charts:
```{r}
barplot(d2$Attack)
 barplot(head(d2$Defence))
 
 
barplot(head(d2$Attack))
title(xlab = 'variable', ylab = 'value')

barplot(head(d2$Attack), xlab = 'variable', ylab = 'value', ylim = c(0,500),col = 'lightblue')
 abline(h = seq(1,9,2), lty = 2, lwd = 0.5, col = 'gray40')
 box()
 
 table(head(d2$Defence))
 
barplot(table(head(d2$Attack)), ylab = 'variable', xlab = 'value')
abline(h = 0)

barplot(head(d2$Attack, beside = TRUE, ylab = 'variable', xlab = 'value'))

barplot(t(head(d2$Attack)), beside = TRUE, legend = TRUE, cex.names = 0.8,col = c('black', 'pink', 'lightblue', 'tan', 'red', 'brown'))
title(ylab = 'variable', xlab = 'value')

bccol = c('black', 'pink', 'lightblue', 'tan', 'red', 'brown')
barplot(head(d2$Attack), beside = TRUE, legend = TRUE, horiz = TRUE,xlim = c(0, 60), col = bccol)
title(ylab = 'variable', xlab = 'vaule')


```

Analysis of variance (ANOVA):
```{r}
bfs.aov = aov(d2$Attack ~ d2$Defence, data = d2)
bfs.aov

summary(bfs.aov)

model.tables(bfs.aov, type = 'means')

pw.aov = aov(d2$Attack ~ d2$Defence+ d2$Sp_attack, data = d2)
pw.aov = aov(d2$Attack ~d2$Defence * d2$Defence, data = d2)
pw.aov
pw.aov = aov(d2$Attack ~ d2$Defence + d2$Sp_attack + Attack:Defence, data = d2)
pw.aov

summary(pw.aov)

pw.aov = aov(d2$Attack ~ d2$Defence * d2$Sp_attack, data = d2)
boxplot(d2$Attack ~ d2$Defence * d2$Sp_attack,  data = d2, cex.axis =0.9)
title(xlab = 'varaible', ylab = 'value')


```

Interaction plots:
```{r}
interaction.plot(d2$Attack, d2$Defence, d2$Sp_attack)

attach(d2)
interaction.plot(d2$HP, d2$Attack, d2$Defence, type = 'b')
detach(d2)

interaction.plot(d2$HP, d2$Attack, d2$Defence, type = 'b', pch = 3:1)

interaction.plot(d2$HP, d2$Attack, d2$Defence, type = 'b', pch = 3:1, lty = 3:1,
col = c('red', 'blue', 'darkgreen'))

interaction.plot(d2$HP, d2$Attack, d2$Defence, type = 'b', pch = 3:1, fun = median)

interaction.plot(d2$HP, d2$Attack, d2$Defence, type = 'b', pch = 3:1,
xlab = 'variable', ylab = 'value',
main = 'Interaction plot')

head(replications(d2))

d.df=data.frame(d2$HP,d2$Attack)
head(d.df)
```

Summarize using a grouping variable:
```{r}
head(aggregate(d2[2:5], by = list(d2$Attack), FUN = sum))

head(aggregate(d2$Attack ~ d2$Defence, data = d2, FUN = mean))

head(aggregate(cbind(d2$Attack, d2$Defence) ~ d2$Sp_attack, data = d2, FUN = mean))

head(aggregate(. ~ d2$Defence, data = d2, FUN = mean))

pw.agg = aggregate(d2$Attack ~ d2$Defence * d2$Sp_attack * d2$HP, data = d2, FUN = mean)
head(pw.agg)

head(aggregate(d2$Attack ~ . , data = d2, FUN = mean))

head(aggregate(d2$HP ~ 1 , data = d2, FUN = mean))
```

Curvilinear regression:

```{r}
lm(d2$HP ~ d2$Attack, data = d2)

fw.lm = lm(d2$HP ~ d2$Attack, data = d2)
summary(fw.lm)

fw.lm$rank

coef(fw.lm)

 confint(fw.lm)
 
 confint(fw.lm, parm = c('(Intercept)', 'HP'), level = 0.9)
 
head(fitted(fw.lm))

head(residuals(fw.lm))

formula(fw.lm)

fw.lm$cal

plot(d2$Attack, d2$HP)
plot(~ d2$Attack + d2$Defence, data = d2)
plot(d2$Attack ~ d2$Sp_defence, data = d2)
abline(lm(d2$HP ~ d2$Attack, data = d2))
abline(a = coef(fw.lm[1], b = coef(fw.lm[2])))
abline(coef(fw.lm))

anova(fw.lm)


mf.lm = lm(d2$HP ~ d2$Attack, data = d2)
mf.lm

pb.lm = lm(d2$HP ~ d2$Attack, data = d2)
pb.lm

mf.lm = lm(d2$HP ~ 1, data = d2)
pb.lm = lm(d2$HP ~ 1, data = d2)
mf.lm
pb.lm

add1(mf.lm, scope = d2)

mf.lm = lm(d2$HP ~ d2$Attack, data = d2)
mf.lm

mf.lm = lm(d2$HP ~ d2$Attack + d2$Defence, data = d2)
summary(mf.lm)

mf.lm = lm(d2$HP ~ d2$Attack, data = d2)
mf.lm

mf.lm1 = lm(d2$Attack ~ d2$HP, data = d2)
mf.lm2 = lm(d2$Attack ~ ., data = d2)
anova(mf.lm1,mf.lm2)

pg.lm = lm(d2$HP ~ log(d2$Attack), data = d2)
summary(pg.lm)

 bbel.lm = lm(d2$HP ~ d2$Attack + I(Attack^2), data = d2)
 bbel.lm

plot(d2$Attack ~ d2$Defence, data = d2)
plot(d2$Attack ~ d2$Sp_attack, data = d2)
lines(d2$Attack, fitted(pg.lm))
lines(d2$Defence, fitted(bbel.lm))

head(predict(mf.lm))

prd = predict(mf.lm, interval = 'confidence', level = 0.95)

attach(d2)
prd = data.frame(prd, d2)
detach(d2)

prd = data.frame(prd)
prd$d2 = d2$d2
head(prd)

plot(d2$HP ~ d2$Attack, data =d2) # basic plot
lines(prd$HP, prd$fit) # also best fit
lines(prd$Attack, prd$lwr, lty = 2) # lower CI
lines(prd$Defence, prd$upr, lty = 2) # upper CI

mf.lm = lm(d2$Attack ~ d2$Defence + d2$Sp_attack, data = d2)
plot(mf.lm)

bf.m = apply(d2, 2, mean, na.rm = TRUE)

bf.sd = apply(d2, 2, sd, na.rm = TRUE)
```

Bar Graphs Using ggplot() with geom_col()

```{r}
library(gcookbook)
library(dplyr)
library(ggplot2)
ggplot(d2, aes(x = d2$Defence, y = d2$Sp_attack)) + geom_col()

ggplot(d2, aes(x = d2$HP, y = d2$Attack)) + geom_col(fill = "yellow", colour = "black")

ggplot(d2, aes(x = d2$HP, y =d2$Attack, fill = HP)) + geom_col(position = "dodge")
```

Bar Graphs Using ggplot() with geom_bar()
```{r}
ggplot(d2, aes(x = Attack)) + geom_bar()
```

Using Colors in a Bar Graph:
```{r}
a2 <- d2 %>% arrange(desc(Attack)) %>% slice(1:10)
a2

```

Adjusting Bar Width and Spacing:

```{r}
ggplot(d2, aes(x = Attack, y = Defence)) + geom_col(width = 0.5)

ggplot(d2, aes(x = Attack, y = HP)) + geom_col(width = 1)

ggplot(d2, aes(x = Attack, y = HP, fill = Attack)) + geom_col(width = 0.5, position = "dodge")

ggplot(d2, aes(x = Attack, y =HP, fill = Attack)) +geom_col(width = 0.5, position = position_dodge(0.7))

ggplot(d2, aes(x = Attack, y = HP, fill = Attack)) +geom_col() + guides(fill = guide_legend(reverse = TRUE))

ggplot(d2, aes(x = Attack, y = HP, fill = Attack)) + geom_col(position = position_stack(reverse = TRUE)) + guides(fill = guide_legend(reverse = TRUE))

ggplot(d2, aes(x = Attack, y = HP, fill = Attack)) + geom_col(position = "fill") + scale_y_continuous(labels = scales::percent)

# using group_by command:
a2 <- d2 %>%group_by(Attack) %>% mutate(Sp_attack = Attack / sum(Attack) * 100)
head(a2)

ggplot(d2, aes(x = Attack, y =Sp_attack , fill = Attack)) + geom_col()

ggplot(d2, aes(x = interaction(Attack, HP), y = Defence)) +geom_col() + geom_text(aes(label = Defence), vjust = 1.5, colour = "white")

```

